Reducing Horton-Strahler Stream Order Can Enhance Flood Inundation Mapping Skill with Applications for the U.S. National Water Model

### 2021-05-18 manuscripts/owp_fim3_2020/sections/introduction.tex
Updates to figures

In 2017 and 2019, river and flash flooding combined represented the leading cause of death among all natural disasters and second in 2018 \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018}. 

@@

One of these operational systems is the Advanced Hydrologic Prediction System (AHPS) maintained by National Oceanic Atmospheric Administration (NOAA) National Weather Service (NWS) with approximately 3,781 forecast points across the US at typically short forecast horizons of 24 or 72 hours \cite{mcenery2005noaa}.
AHPS provides forecasting services in the form of ensemble stream flows at 3,781 forecast points illustrated in Figure \ref{fig:all_ahps_points} and flood inundation maps (FIM) at 188 of those forecast points shown in Figure \ref{fig:fim_ahps_points}.
\note{Fernando A}{Combine Figure 1 and 2 with one figure: All AHPS forecast points with no inundation extents one color and with extents other color for entire US if needed}

@@

On an approximate basis, there is only one forecast point every 1,455 km of river and one forecast point with FIM every 29,261 km.

@@

\includegraphics[scale=2.0]{figures/ahps_all_forecast_points.jpg}
\caption{All 3,781 forecast points in United States' Advanced Hydrologic Prediction System for lower 48 states and Alaska. No forecast points in Hawaii or territories. From: water.weather.gov/ahps}
\label{fig:all_ahps_points}
\includegraphics[scale=2.0]{figures/ahps_fim_forecast_points_conus.jpg}
\caption{All 188 forecast points in United States' Advanced Hydrologic Prediction System with flood inundation mapping capabilities. One forecast point near Juneau, Alaska not shown. No forecast points in Hawaii or remaining territories. From: water.weather.gov/ahps}
\label{fig:fim_ahps_points}

@@

@@

@@

\citeA{zheng2018river} developed methodology for determine stage-discharge relationships known as synthetic rating curves (SRC) by sampling reach-averaged parameters from HAND datasets and inputting into the Manning's equation \cite{gauckler1867etudes,manning1890flow}.

@@

More notably the software added the ability to derive drainage potentials on a per mainstem basis as originally conceptualized by \citeA{mcgehee2016modified}.
Mainstem is generally used here as defined in \citeA{blodgett2020mainstems} but only available in mainstems downstream of forecast AHPS points within the workflow introduced by \citeA{djokic2019arc}.

@@

These findings suggest improvements to HAND are required that utilize advanced computational algorithms and software to compute a FIM hydrofabric required for producing continental-scale FIM (CFIM).

>>>>

In 2017 and 2019, river and flash flooding combined represented the leading cause of death and the second leading cause in 2018 among all natural disasters in the US \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018}. 

@@

One of these operational systems is the Advanced Hydrologic Prediction System (AHPS) maintained by National Oceanic Atmospheric Administration (NOAA) National Weather Service (NWS) with more than 3,000 forecast points across the US at typically short forecast horizons of 24 or 72 hours \cite{mcenery2005noaa}.
AHPS provides forecasting services in the form of ensemble stream flows at more than 3,000 forecast points and flood inundation maps (FIM) at more than 150 of those locations shown in Figure \ref{fig:forecast_points}.
Besides the AHPS forecast points, we display in Figure \ref{fig:forecast_points} two forecasting networks relevant to the National Water Model (NWM) which will be introduced in Section \ref{ssec:national_water_model}.

@@

On an approximate basis, there is only one forecast point every 1,450 km of river and one forecast point with FIM every 29,000 km.

@@

\includegraphics[scale=1.0]{figures/forecast_points.jpg}
\caption{Forecast points with and without FIM in United States' Advanced Hydrologic Prediction System. Also show are the National Water Model stream networks at the full resolution (FR) and Replace and Reroute (RnR) Mainstems (MS) resolution.}
\label{fig:forecast_points}

@@

\label{ssec:national_water_model}

@@

Figure \ref{ssec:national_water_model} shows Full-Resolution (FR) modeling stream network as well as the RnR mainstems (MS) network.

@@

\citeA{zheng2018river} developed methodology for determining stage-discharge relationships known as synthetic rating curves (SRC) by sampling reach-averaged parameters from HAND datasets and inputting into the Manning's equation \cite{gauckler1867etudes,manning1890flow}.

@@

More notably the software added the ability to derive drainage potentials on both the NWM FR and MS stream networks which leverages the lower drainage density and Horton-Strahler stream order of the MS network.

@@

These findings suggest improvements to HAND are required that utilize advanced computational algorithms and software to compute a FIM hydrofabric required for producing continental-scale FIM.

### 2021-05-18 manuscripts/owp_fim3_2020/sections/keypoints.tex
Updates to figures

\begin{keypoints}
\item KP1
\item KP2
\item KP3
\end{keypoints}

>>>>

\begin{keypoints}
\item Height Above Nearest Drainage (HAND) is a terrain index used for rapid generation of flood inundation maps (FIM) across large scales.
\item Despite its advantages, FIM generated from HAND suffers from a number of limitations most notably its dependence on drainage density.
\item Among other developments, here we present a means of resolving stream density by reducing drainage density to a singular flow path. 
\end{keypoints}
